---
title: "Strings and factors"
author: "Ravi Brenner"
date: "2024-10-15"
output: github_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)
```

## Strings

```{r}
string_vec <- c("my", "name", "is", "Ravi")
```

```{r}
str_detect(string_vec, "a")
str_detect(string_vec, "Ravi")
str_detect(string_vec, " Ravi")
str_detect(string_vec, "ravi")
```

```{r}
str_replace(string_vec, "Ravi", "ravi")
str_replace(string_vec, "a", "A")
```

Those were basic/easy. Finding exact matches

Now it gets more complicated
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )
```

```{r}
str_detect(string_vec, "i think")

# line starts with / beginning of the line
str_detect(string_vec, "^i think")

# line ends with / end of the line
str_detect(string_vec, "i think$")
```


```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )
```

```{r}
str_detect(string_vec, "pumpkin")
str_detect(string_vec, "Pumpkin")
str_detect(string_vec, "PUMPKIN")

# want multiple options
str_detect(string_vec, "[Pp]umpkin")
```

Can work with numbers and letters in specific arrangement

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )
```

Want: number followed by two letters

This is saying: any number 0-9, followed by and number a-z (case sensitive)
```{r}
str_detect(string_vec, "[0-9][a-z]")
```

```{r}
str_detect(string_vec, "[0-9][a-zA-Z]")
```

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )
```

can use '.' to match any character
```{r}
str_detect(string_vec, "7.11")
```

Then things can get really strange

use "\" to use a special character. Since "\" is itself a special character, we need to put "\\".
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```








