---
title: "Strings and factors"
author: "Ravi Brenner"
date: "2024-10-15"
output: github_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)
```

## Strings

```{r}
string_vec <- c("my", "name", "is", "Ravi")
```

```{r}
str_detect(string_vec, "a")
str_detect(string_vec, "Ravi")
str_detect(string_vec, " Ravi")
str_detect(string_vec, "ravi")
```

```{r}
str_replace(string_vec, "Ravi", "ravi")
str_replace(string_vec, "a", "A")
```

Those were basic/easy. Finding exact matches

Now it gets more complicated
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )
```

```{r}
str_detect(string_vec, "i think")

# line starts with / beginning of the line
str_detect(string_vec, "^i think")

# line ends with / end of the line
str_detect(string_vec, "i think$")
```


```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )
```

```{r}
str_detect(string_vec, "pumpkin")
str_detect(string_vec, "Pumpkin")
str_detect(string_vec, "PUMPKIN")

# want multiple options
str_detect(string_vec, "[Pp]umpkin")
```

Can work with numbers and letters in specific arrangement

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )
```

Want: number followed by two letters

This is saying: any number 0-9, followed by and number a-z (case sensitive)
```{r}
str_detect(string_vec, "[0-9][a-z]")
```

```{r}
str_detect(string_vec, "[0-9][a-zA-Z]")
```

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )
```

can use '.' to match any character
```{r}
str_detect(string_vec, "7.11")
```

Then things can get really strange

use "\" to use a special character. Since "\" is itself a special character, we need to put "\\".
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

## Factors
Apparently people hate these

```{r}
sex_vec <- factor(c("male", "male", "female", "female"))

as.numeric(sex_vec)
```

Relevel
```{r}
sex_vec = fct_relevel(sex_vec, "male")
```

## Revisit some examples from the past

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)

marj_use_table <- drug_use_html |> 
  html_table() |> 
  first() |>
  slice(-1)
```

really need to clean and tidy this table if we want to use it
```{r}
marj_use_df <- marj_use_table |>
  select(-contains("P Value")) |>
  pivot_longer(cols  = -State,
               names_to = "age_year",
               values_to = "percent") |>
  separate(age_year,
           into  = c("age","year"),
           sep = "\\(") |>
  mutate(
    year = str_remove(year,"\\)"),
    percent = str_remove(percent,"[a-c]$"),
    percent = as.numeric(percent)) |>
  filter(!State %in% c("Total U.S.","Northeast","Midwest","South","West"))
```

```{r}
marj_use_df |>
  filter(age == "12-17") |>
  mutate(State = fct_reorder(State,percent)) |>
  ggplot(aes(x = State,y = percent,color = year)) +
  geom_point() + 
  scale_color_viridis_d() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## NYC restaurant inspections
```{r}
data("rest_inspec")
```

```{r}
rest_inspec |>
  count(boro, grade) |>
  pivot_wider(names_from = grade, values_from = n)
```

```{r}
rest_inspec <- rest_inspec |>
  filter(str_detect(grade,"[A-C]"),
         ! boro == "Missing")
```

```{r}
rest_inspec |>
  mutate(dba = str_to_sentence(dba)) |>
  filter(str_detect(dba, "Pizza")) |>
  mutate(boro = fct_relevel(boro, "STATEN ISLAND")) |>
  ggplot(aes(x = boro)) + 
  geom_bar()
```


```{r}
rest_inspec |>
  mutate(dba = str_to_sentence(dba)) |>
  filter(str_detect(dba, "Pizza")) |>
  mutate(boro = fct_infreq(boro),
         boro = fct_recode(boro,"THE CITY" = "MANHATTAN")) |>
  ggplot(aes(x = boro)) + 
  geom_bar()
```

One last thing on factors:
```{r}
rest_inspec |>
  mutate(dba = str_to_sentence(dba)) |>
  filter(str_detect(dba, "Pizza")) |>
  mutate(boro = fct_infreq(boro)) |>
  lm(zipcode ~ boro, data = _) 
```

