---
title: "Reading data from the web"
author: "Ravi Brenner"
date: "2024-10-10"
output: github_document
---

```{r setup}
library(tidyverse)
library(rvest)
library(httr)
```

## Web scraping basics

Load data from website with tables on it
```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)

drug_use_html
```

Attempt to get pieces I actually need:
```{r}
marj_use_df = 
  drug_use_html |> 
  html_table() |> 
  first() |>
  slice(-1)
```

Learning assessment: Create a data frame that contains the cost of living table for New York from this page.
https://www.bestplaces.net/cost_of_living/city/new_york/new_york

```{r}
read_html("https://www.bestplaces.net/cost_of_living/city/new_york/new_york") |>
  html_table(header = TRUE) |>
  first()
```

Needs to be tidied, but we have the data

## CSS Selectors
```{r}
swm_url = "https://www.imdb.com/list/ls070150896/"

swm_html <- read_html(swm_url)
```

How to figure out data we actually want?
```{r}
swm_title_vec <- 
  swm_html |>
  html_elements(".ipc-title-link-wrapper .ipc-title__text") |>
  html_text()

swm_runtime_vec <- 
  swm_html |>
  html_elements(".dli-title-metadata-item:nth-child(2)") |>
  html_text()

swm_score_vec <- 
  swm_html |>
  html_elements(".metacritic-score-box") |>
  html_text()

swm_df <- tibble(title = swm_title_vec,
                 runtime = swm_runtime_vec,
                 score = swm_score_vec)
```

Trying with another website!
```{r}
books_url = "https://books.toscrape.com/"

books_html = read_html(books_url)
```

```{r}
books_html |>
  html_elements(".product_pod a") |>
  html_text2()
```

## Using APIs

Get water data from NYC open data
```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") |> 
  content()
```

BRFSS data
```{r}
brfss_df <-
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000)) |>
  content()
```

